%$Id: resume.tex,v 1.2 2011/12/16 04:01:44 raman Exp $
\documentclass{article}
\input{resume-style}
\title{Curriculum Vit\ae}
\author{T. V. Raman}

\begin{document}
\begin{center}
\begin{tabular}{||c||}
{T.\  V.\  Raman}\newrow
{Google Research}\newrow
(650) 799-5724 \newrow
\href{mailto:tv.raman.tv@gmail.com}{\email{tv.raman.tv@gmail.com}}  \newrow
\livelink{http://emacspeak.sf.net/raman}\newrow
\end{tabular}
\end{center}

\section*{Summary}

I am an accomplished Computer Scientist with over $17$~years of
leadership experience in advanced technology development. During
this time, I have authored $3$~books and filed over $50$~patents;
my work on auditory interfaces has been profiled in mainstream
publications including the
\href{http://www.nytimes.com/2009/01/04/business/04blind.html?_r=1}{New
  York Times} and
\href{http://emacspeak.sourceforge.net/raman/sciam-0996profile.html}{Scientific
  American}. I presently lead accessibility for
\href{http://eyes-free.googlecode.com}{Google Android} and
\href{http://google-axs-chrome.googlecode.com}{Google Chrome}; I
have leading edge expertise in developing auditory interfaces for
mobile devices and Web applications. Earlier, I led the
definition of numerous W3C standards including XForms and Aural
CSS.

\section*{Objective} 

Deliver technologies that enable ubiquitous, eyes-free access to
the emerging Web platform when accessed from a wide variety of
devices ranging from smart phones and tablets to network-based
computers. Speech is the next  dimension in user
interfaces, and I am developing application frameworks that
combine speech technologies with the power of the Web Cloud to
deliver innovative solutions that enable anytime, anywhere
access.

\section*{Work experience}
\begin{itemize}
  \item {Google},
  \href{http://www.google.com/search?q=\%22tv+raman\%22}{Google Research}, Mountain View, CA\\
  \textbf{Research Scientist}.
\hfill Aug~2005--Present.
\begin{description}
\item[\href{http://eyes-free.googlecode.com}{Android Access}]  Led Android Accessibility from its
inception to deliver many innovative end-user solutions.
\item[\href{http://google-axs-chrome.googlecode.com}{ChromeVox}]  Led the design of a  complete accessibility
solution that is built entirely of Web technologies.
\item[\href{http://labs.google.com/accessible}{Accessible Search}] Built an innovative classifier for
  measuring Web Accessibility  that is integrated into Google.
  
\end{description}

\item {IBM Research},
\href{http://www.google.com/search?q=tv+raman+IBM+Almaden}{Almaden Research Center}, San Jose, CA\\
  \textbf{Research Staff Member}: Architect, Conversational
  Multimodal WWW\@. \hfill Aug~1999--Aug~2005.
\begin{description}
\item [\href{http://www.amazon.com/XForms-XML-Powered-Web-Forms/dp/0321154991}{XForms}] Authoring applications for the next generation
  WWW.
\item [RDC] Reusable Dialog Components to speech-enable the Web.

\item[X$+$V] Speech-enabling XHTML to create a \emph{multimodal}
  Web.
\end{description}
\item   {Adobe Systems},  {Advanced Technology Group}, San Jose, CA\\
  \textbf{Senior Computer Scientist}: Dynamic publishing on the
  Internet.  \hfill Oct~1995--Aug~1999.
\begin{description}
\item[PDF2HTML] Developed the PDF to HTML translator bundled with
  major Web search engines \mdash
  \href{http://access.adobe.com}{access.adobe.com}.
  
\item[XML Metadata] Developed an XML-based virtual document
  architecture to enable cross-application content reuse.
\end{description}
\item   {Digital Equipment Corporation},  {Cambridge Research Lab}, Cambridge, MA\\
  \textbf{Research Staff}: \textbf{Retriever} --A Multimodal Web
  Interface.  \hfill Feb~1994--Oct~1995.
\item {Intel Corporation}, {Intel Architecture Labs},
  Hillsboro, OR\\
  \textbf{Summer Associate}: Prototyped an email telephony
  interface. \hfill Jun--Aug~1993.
\item {Xerox Palo Alto Research Center}, Palo Alto, CA\\
  \textbf{Summer Associate}: Prototyped a new reading machine
  architecture. \hfill May--Aug~1991.
\end{itemize}

\longpage

\section*{Education}

\begin{itemize}
\item \textbf{Cornell University}, Ithaca, NY
\begin{itemize}
\item \textbf{PhD.\ Applied Mathematics}: \hfill
  Aug~1989--Jan~1994.\\
  \hspace*{.1in} Awarded the
\href{http://www.acm.org/awards/dd_citation/1994B.html}{%
  ACM Doctoral Dissertation Award, 1994}.\\
\hspace*{.1in} {Thesis}: {Audio System For Technical Readings}.
 {Adviser}: Prof. David Gries, Computer Science.
\item \textbf{MS Computer Science}: \hfill May~1992.
\end{itemize}
\item \textbf{Indian Institute of Technology}, Bombay, India:
  \textbf{MSc Computer Science}: \textit{GPA}:~$9.78/10.00$
  \hfill July~1989.
\item \textbf{University of Pune}, Pune, India: \textbf{BA
    Mathematics}: \hfill May~1987.
\end{itemize}


\section*{Selected Awards and Honors}
\begin{itemize}
\item \textbf{Computerworld Award} {Smithsonian Institution}
  \href{http://emacspeak.sf.net}{Emacspeak:} Complete Audio
  Desktop. \hfill April~1999.
\item \textbf{Association of Computing Machinery (ACM)}
  \href{http://www.acm.org/awards/dd_citation/1994B.html}{%
    Doctoral Dissertation Award} \hfill1994.
\item \textbf{Intel Graduate Fellowship} {Intel Corporation, CA}
  \hfill 1992.
\item \textbf{Graduate Fellowship} Cornell University.  \hfill
  1989.
\item \textbf{President's Silver Medal} {\iitb}.  \hfill 1989.
\item \textbf{Sir Cusrow Wadia Gold Medal} {University of Pune}.
  \hfill 1987.
\item \textbf{Sir Ness Wadia Gold Medal}.  \hfill 1984.
  \vspace*{-6pt}
\end{itemize}

\section*{Selected Books,\ Patents And Software}
%warning --remember to escape % in urls 

\begin{tabular}{||c|c|c|c||}\hline
\multicolumn{4}{|c|}{Locating My Publications}\newrow
\href{http://scholar.google.com/citations?user=mVNjyD8AAAAJ}{Google Scholar}&
        \href{http://portal.acm.org/results.cfm?query=T.\%2BV.\%2BRaman}{ACM}&
        \href{http://liinwww.ira.uka.de/searchbib/index?query=RamanT;partial=on;case=on;results=citation;maxnum=200}{CSB}&
\href{http://www.informatik.uni-trier.de/~ley/db/indices/a-tree/r/Raman:T=_V=.html}{DBLP}\newrow
  \end{tabular}

\begin{plainbibliography}{}
\bibitem{raman:xf03} T.~V. Raman.  
\newblock {\em XForms --- XML
    Powered Web Forms}.
\newblock \href{http://www.amazon.com/exec/obidos/ASIN/0321154991}%
{Addison Wesley},  2003.
\bibitem{raman:lncs98} T.~V. Raman.  
\newblock {\em Audio System
    For Technical Readings}.  
\newblock
\href{http://link.springer-ny.com/link/service/series/0558/tocs/t1410.htm}%
{LNCS 1410}, Springer Verlag,  1998.
\bibitem{raman:aui97} T.~V. Raman.  
\newblock {\em Auditory User Interfaces}.
\newblock \href{http://www.cs.cornell.edu/home/raman/aui/aui.html}%
{Kluwer Academic Publishers},  1997.
\bibitem{raman:ip-aster96} T.~V. Raman.  
\newblock
  \emph{Generating audio renderings of digitized works}.
\newblock Cornell.
\newblock \href{http://patft.uspto.gov/netacgi/nph-Parser?TERM1=5,572,625&Sect1=PTO1&Sect2=HITOFF&d=PALL&p=1&u=\%2Fnetahtml\%2Fsrchnum.htm&r=0&f=S&l=50}%
{U.S.\ Patent 5,572,625},  1996.
\bibitem{raman:ip-intel93} T.\ V.\ Raman and Jim Larson.
\newblock \emph{Telephone access system}.  
\newblock Intel Corporation.
\newblock \href{http://patft.uspto.gov/netacgi/nph-Parser?TERM1=5,825,854&Sect1=PTO1&Sect2=HITOFF&d=PALL&p=1&u=\%2Fnetahtml\%2Fsrchnum.htm&r=0&f=S&l=50}%
{U.S.\ Patent 5,825,854},  1998.
\bibitem{raman:ip-retriever98} T.~V. Raman.  
\newblock
  \emph{Multimodal information presentation system}.  
\newblock DEC.
\newblock \href{http://patft.uspto.gov/netacgi/nph-Parser?TERM1=5,748,186&Sect1=PTO1&Sect2=HITOFF&d=PALL&p=1&u=\%2Fnetahtml\%2Fsrchnum.htm&r=0&f=S&l=50}%
{U.S.\ Patent 5,748,186,}  1998.
\bibitem{raman:ip-proxy97} T.~V. Raman.  
\newblock \emph{Data
    stream processing on networks}.  
\newblock Adobe Systems.
\newblock \href{http://patft.uspto.gov/netacgi/nph-Parser?TERM1=6,134,598&Sect1=PTO1&Sect2=HITOFF&d=PALL&p=1&u=\%2Fnetahtml\%2Fsrchnum.htm&r=0&f=S&l=50}%
{U.S.\ Patent 6,134,598},  2000.
\bibitem{raman:ip-capture98} T.~V.\ Raman and John~Warnock.
\newblock \emph{Digitized speech and text}.  
\newblock Adobe
  Systems.
\newblock \href{http://patft.uspto.gov/netacgi/nph-Parser?TERM1=6,151,576&Sect1=PTO1&Sect2=HITOFF&d=PALL&p=1&u=\%2Fnetahtml\%2Fsrchnum.htm&r=0&f=S&l=50}%
{U.S.\ Patent 6,151,576},  2000.
\bibitem{raman:ip-ddf97} T.~V. Raman.  
\newblock \emph{Document
    description format}.  
\newblock Adobe Systems.
\newblock \href{http://patft.uspto.gov/netacgi/nph-Parser?TERM1=6,249,794&Sect1=PTO1&Sect2=HITOFF&d=PALL&p=1&u=\%2Fnetahtml\%2Fsrchnum.htm&r=0&f=S&l=50}%
{U.S.\ Patent 6,249,794}, , 2001.
\bibitem{raman:ip-emacspeak95} T.~V. Raman.  
\newblock
  \emph{Speech interface for computer application programs}
\newblock DEC.  
\newblock
  \href{http://patft.uspto.gov/netacgi/nph-Parser?Sect1=PTO1&Sect2=HITOFF&d=PALL&p=1&u=/netahtml/srchnum.htm&r=1&f=G&l=50&s1='6289312'.WKU.&OS=PN/6289312&RS=PN/6289312}
  {U.S.\ Patent 6,289,312},  2001.
\bibitem{raman:ip-mmi05} T.~V. Raman.  et al 
\newblock
  \emph{Dialog management  in a
multimodal environment}
\newblock IBM.  
\newblock
  \href{http://patft.uspto.gov/netacgi/nph-Parser?Sect1=PTO1\&Sect2=HITOFF\&d=PALL\&p=1\&u=/netahtml/srchnum.htm\&r=1\&f=G\&l=50\&s1=6839896.WKU.\&OS=PN/6839896\&RS=PN/6839896}
  {U.S.\ Patent 6,839,896},  2005.
\bibitem{raman:ip-access05} T.~V. Raman.  et al 
\newblock
  \emph{Web accessibility service apparatus and method}
\newblock IBM.  
\newblock
  \href{http://patft.uspto.gov/netacgi/nph-Parser?Sect1=PTO1\&Sect2=HITOFF\&d=PALL\&p=1\&u=/netahtml/srchnum.htm\&r=1\&f=G\&l=50\&s1=6839896.WKU.\&OS=PN/6922726\&RS=PN/6839896}
  {U.S.\ Patent 6,922,726},  2005.
\bibitem{raman:emacspeak-sf} T.~V. Raman.  
\newblock {\em
Emacspeak --- The Complete Audio Desktop}.
\newblock \href{http://emacspeak.sf.net}%
{Open Source Software},  1995.
\end{plainbibliography}

\section*{Selected Publications And Articles}

\begin{plainbibliography}{}
\bibitem{raman2009toward}
TV~Raman.
\newblock Toward 2 w, beyond web 2.0.
\newblock \href{http://emacspeak.sourceforge.net/raman/publications/beyond-web20-cacm-2009/}{\em Communications of the ACM}, 52(2):52--59, 2009.
\bibitem{raman:sciam97} T.~V. Raman.  
\newblock Netsurfing without a monitor.  
\newblock {\em Scientific American}, March 1997.
\newblock \href{http://www.sciam.com/0397issue/0397raman.html}%
{Special Internet Edition.}
\bibitem{raman:ddj97} T.~V.  Raman.
\newblock \href{http://www.ddj.com/ddj/1997/1997.09/rama.htm}%
{Emacspeak ---a speech enabling interface.}  
\newblock {\em Dr.\ 
  Dobb's Journal}, September 1997.
\bibitem{raman:ddj-oped97} T.~V.  Raman.  
\newblock
\href{http://www.ddj.com/editorial/1997/raman.htm}%
{User interface ---a means to an end.}  
\newblock {\em Dr.\ 
  Dobb's Journal}, August 1997.
\bibitem{gibbs:sciam96} Wayt Gibbs.  
\newblock
\href{http://www.sciam.com/0996issue/0996profile.html}%
{Profile: T.\ V.\ raman: Envisioning speech.}  
\newblock {\em
  Scientific American}, September 1996.
\bibitem{hayes:amsci96} Brian Hayes.
\newblock  \href{http://www.amsci.org/amsci/issues/Comsci96/compsci96-03.html}%
{Speaking of mathematics.}  
\newblock {\em American Scientist},
84(2), March--April 1996.
\bibitem{raman:acss97} T.~V. Raman.  
\newblock Cascaded speech
  style sheets.  
\newblock {\em WWW6 Conference, CA.}, April
  1997.
\bibitem{raman:phd-thesis} T.~V.  Raman.  
\newblock
\href{http://www.cs.cornell.edu/home/raman/aster/aster-toplevel.html}%
{{\em Audio System for Technical Readings}.}  
\newblock PhD
thesis, Cornell University, May 1994.
\bibitem{raman:chi96} T.~V.  Raman.  
\newblock Emacspeak --a
  speech interface.  
\newblock {\em CHI96}, April 1996.
\bibitem{2001:w3c-xforms} T.~V.  Raman et al.  
\newblock XForms
  1.0 
\newblock \emph{W3c}, October, 2003.  
\newblock
  \livelink{http://www.w3.org/tr/xforms}
\bibitem{2003:w3c-xml-events} T.~V.  Raman et al.  
\newblock 
  XML Events 
\newblock \emph{W3c}, 2003.  
\newblock
  \livelink{http://www.w3.org/tr/xml-events}
\bibitem{2001:w3c-xhtml-voice} T.~V.  Raman et al.  
\newblock
  {Adding Spoken Interaction To XHTML} 
\newblock \emph{W3c},
  December, 2001.  
\bibitem{xmlj03:xforms} T.~V.  Raman 
\newblock
  \href{http://www.sys-con.com/xml/article.cfm?id=578}%
{Collecting Business Critical Information Using XForms} 
\newblock
  \emph{XML Journal}, April, 2003.
\end{plainbibliography}

\begin{center}
  \end{center}

\section*{Other Interests}
My favorite hobby is
\href{http://www.cs.cornell.edu/home/raman/publications/polyhedra}%
{recreational mathematics}.  I enjoy working on puzzles,
especially those that involve an intuitive feel for mathematics.
One of the things I enjoyed doing the most in the early eighties
was to solve the Rubik's~cube faster than anyone else around~me,
on an average of~about 
\href{http://video.google.com/videoplay?docid=-4180435763269825467}{thirty~seconds}!  During the last few
years, discovering
\href{http://emacspeak.sf.net/raman/publications/polyhedra/}{Zome
  Systems} for building complex polyhedra has helped rekindle my
interest in polyhedral geometry.  I am also interested in
linguistics and can speak about eight~languages, including
French, German and several Indian~languages.
% \section*{References}

% \begin{tabular}{||l|l|l||}\hline
% Prof.\ David Gries &\textit{gries@cs.cornell.edu} &
% \textit{607 255 9207}\newrow
% Prof.\ John E.\ Hopcroft &\textit{jeh@cs.cornell.edu} &
% \textit{607 255 7416}\newrow
% \end{tabular}

\end{document}

